---
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- badges: start -->
![](https://img.shields.io/github/r-package/v/JesseAlderliesten/progutils?color=blue)
[![R-CMD-check](https://github.com/JesseAlderliesten/progutils/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JesseAlderliesten/progutils/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->


# progutils
`progutils` contains utility functions, for example to reorder values, check
equality, construct messages, or handle files and directories.

## Folder structure
```
├── .github
│   └── workflows: workflows to run tests with GitHub Actions
├── R: functions
├── inst
│   └── tinytest: tests
├── man: help-files
└── tests: setup to use 'tinytest' for testing
```

## Installation
You can install `progutils` from 
[GitHub](https://github.com/JesseAlderliesten/progutils) with:
    
```{r, eval = FALSE}
if(!requireNamespace("remotes", quietly = TRUE)) {
  install.packages(pkgs = "remotes", quiet = FALSE)
}
remotes::install_github(repo = "JesseAlderliesten/progutils", dependencies = TRUE,
                        upgrade = FALSE, force = FALSE, quiet = FALSE,
                        build_vignettes = TRUE, lib = NULL,
                        verbose = getOption("verbose"))
```

For information about installing and configuring R and RStudio, see my
repository [checkrpkgs](https://github.com/JesseAlderliesten/checkrpkgs).


## Examples
Every once in a while, reading data into R results in a factor that should be
converted to a numeric vector. The code below illustrates that using
`as.numeric(x)` does *not* work to achieve this while `as.numeric_safe()`
*does* work.

```{r as.numeric_safe}
library(progutils)
x <- factor(11:13)
x
as.numeric(x) # Returns the indices instead of the values!
as.numeric_safe(x) # Returns the values.
```

Using `relevel` or `levels()` to reorder factor levels does *not* work, whereas
function `reorder_levels()` *does*:

```{r reorder_levels}
library(progutils)
f_orig <- factor(letters[c(12:13, 13:11)], levels = letters[13:11])
f_orig

f_test <- f_orig
levels(f_test) <- letters[11:13]
f_test # Also changes the values!
reorder_levels(x = f_orig, new_order = letters[11:13]) # Only changes the levels
```

Using a named numeric vector in the subtitle of a plot requires converting it a
to a character string wrapped to an appropriate length, to prevent text from
running of the plot. Combining `vect_to_char()` and `wrap_text()` provides a way
to achieve that:

```{r wrap_text}
library(progutils)
msg_part_one <- "Some\nmessage you want to display"
num_vect <- c(j = 10, k = 11, l = 12)

cat(wrap_text(paste(msg_part_one, vect_to_char(num_vect)), width = 20))

# Retaining newlines that were present in the original message:
cat(wrap_text(paste(msg_part_one, vect_to_char(num_vect)), width = 20,
              ignore_newlines = FALSE))
```


# Alternatives
Many packages with miscellaneous functions exist on [GitHub](https://github.com/)
and [CRAN](https://cran.r-project.org/web/packages/available_packages_by_name.html):
frequently they have 'misc' or 'util' in their title or description. In addition,
many packages have an `utils` folder in `R`.
